\bf Eye Fitting Straight Lines in the Modern Era

### 2021-09-09 Eye-Fitting-Stright-Lines-in-the-Modern-Era.tex
rough rough rough draft results for first part...

\newcommand{\ear}[1]{{\textcolor{blue}{#1}}}
\newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
\newcommand{\rh}[1]{{\textcolor{Green}{#1}}}
\usepackage[capitalise]{cleveref}
\newcommand\pcref[1]{(\cref{#1})}
\usepackage{algorithm,algpseudocode,booktabs}

@@

Advanced technology and computing power have promoted data visualization
as a central tool in modern data science. \citet{unwin2020data} defines
data visualization as the art of drawing graphical charts in order to
display data. Graphics are useful for data cleaning, exploring data
structure, and have been an essential component in communicating
information for the last 200 years \citep{lewandowsky1989perception}.
Although statistical graphics have become widely used and valued in
science, business, and in many other aspects of life, as creators of
graphics, we are too accepting of them as default without asking
critical questions about the graphics we create or view (Unwin, 2020).

@@

@@

One way in which we evaluate the effectiveness of charts is through the
use of graphical tests. We could ask participants to identify
differences in graphs, read information off of a chart accurately, use
data to make correct real-world decisions, or predict the next few
observations. All of these types of tests require different levels of
use and manipulation of the information being presented in the chart.

\begin{itemize}
\tightlist
\item
  Psychophysics (speed and accuracy)
\item
  Lineups
\end{itemize}

\hypertarget{trend-judgement}{%
\subsection{Trend Judgement}\label{trend-judgement}}

\begin{itemize}
\tightlist
\item
  \citet{finney1951subjective}
\item
  \citet{mosteller_eye_1981}
\item
  \citet{ciccione2021can}
\end{itemize}

Initial studies in the 20th century explored the use of fitting lines by
eye through a set of points
\citep{finney1951subjective, mosteller_eye_1981}. Common methods of
fitting trends by eye involve maneuvering a string, black thread, or
ruler until the fit is suitable, then drawing the line through the set
of points. In \citet{finney1951subjective}, it was of interest to
determine the effect of stopping iterative maximum likelihood
calculations after one iteration. Many techniques in statistical
analysis are performed with the aid of iterative calculations such as
Newton's method or Fisher's scoring. Guesses are made at the best
estimates of certain parameters and these guesses are then used as the
basis of a computation which yields a new set of approximation to the
parameter estimates; this same procedure is the performed on the new
parameter estimates and the computing cycle is repeated until
convergence, as determined by the statistician, is reached. The author
was interested in whether one iteration of calculations was sufficient
in the estimation of parameters connected with dose-response
relationships. One measure of interest is the relative potency between a
test preparation of doses and standard preparation of does; relative
potency is calculated as the ratio of two equally effective doses
between the two preparation methods. \cref{fig:subjective-judgement}
shows a pair of parallel probit responses in a biological assay. The
x-axis is the \(\log_{1.5}\) dose level for four dose levels (for
example, doses 4, 6, 9, and 13 correspond correspond to equally spaced
values on a logarithmic scale, labeled 0, 1, 2, and 3) and the y-axis is
the corresponding probit response as calculated in
\citet{finney1948table}; circles correspond to the test preparation
method while the crosses correspond to the standard preparation method.
For these sort of assays, the does-response relationship follows a
linear regression of the probit response on the logarithm of the dose
levels; the two preparation methods can be constrained to be parallel
\citep{jerne1949validity}, limiting the relative potency to one
consistent value. In this study, twenty-one scientists were recruited
via postal mail and asked to ``rule two lines'' in order to judge by eye
the positions for a pair of parallel probit regression lines in a
biological assay. The author then computed one iterative calculation of
the relative potency based on starting values as indicated by the pair
of lines provided by each participant and compared these relative
potency estimates to that which was estimated by the full probit
technique (reaching convergence through multiple iterations). Results
indicated that one cycle of iterations for calculating the relative
potency was sufficient based on the starting values provided by eye from
the participants.

Thirty years later, Mosteller et al.~(1981), sought to understand the
properties of least squares and other computed lines by establishing one
systematic method of fitting lines by eye. The authors recruited 153
graduate students and post doctoral researchers in Introductory
Biostatistics. Participants were asked to fit lines by eye to four
scatterplots using an 8.5 x 11 inch transparency with a straight line
etched completely across the middle. A latin square design
\citep{giesbrecht2004planning} with packets of the set of points stapled
together in four different sequences was used to determine if there is
an effect of order of presentation. It was found that order of
presentation had no effect and that participants tended to fit the slope
of the first principal component (error minimized orthogonally, both
horizontal and vertical, to the regression line) over the slope of the
least squares regression line (error minimized vertically to the
regression line).

In 2015, the New York Times introduced an interactive feature, called
You Draw It
\citep{aisch_cox_quealy_2015, buchanan_park_pearce_2017, katz_2017}.
Readers are asked to input their own assumptions about various metrics
and compare how these assumptions relate to reality. The New York Times
team utilizes Data Driven Documents (D3) that allows readers to predict
these metrics through the use of drawing a line on their computer screen
with their computer mouse. \citep{katz_2017} is one such example in
which readers are asked to draw the line for the missing years providing
what they estimate to be the number of Americans who have died every
year from car accidents, since 1990. After the reader has completed
drawing the line, the actual observed values are revealed and the reader
may check their estimated knowledge against the actual reported data.

Major news and research organizations such as the New York Times,
FiveThirtyEight, Washington Post, and the Pew Research Center create and
customize graphics with Data Driven Documents (D3). In June 2020, the
New York Times released a front page displaying figures that represent
each of the 100,000 lives lost from the COVID-19 pandemic until this
point in time \citep{NYTrememberinglives}; this visualization was meant
to bring about a visceral reaction and resonate with readers. During
2021 March Madness, FiveThirtyEight created a roster-shuffling machine
which allowed readers to build their own NBA contender through
interactivity \citep{ryanabest_2021}. Data Driven Documents (D3) is an
open-source JavaScript based graphing framework created by Mike Bostock
during his time working on graphics at the New York Times. For readers
familiar with R, it is notable to consider D3 in JavaScript equivalent
to the ggplot2 package in R \citep{ggplot2}.

\hypertarget{research-objective}{%
\subsection{Research objective}\label{research-objective}}

The goal of this paper is to establish you draw it as a tool for
measuring predictions of trends fitted by eye and a method for testing
graphics. In order to validate you draw it as a method for testing
graphics, the first sub-study, referred to as Eye Fitting Straight Lines
in the Modern Era, replicated the experiment and results found in
Mosteller et al.~(1981).

\hypertarget{study-development}{%
\section{Study Development}\label{study-development}}

\hypertarget{interactive-plot-development}{%
\subsection{Interactive Plot
Development}\label{interactive-plot-development}}

@@

All data processing was conducted in R before being passed to the D3.js
source code. A total of \(N = 30\) points \((x_i, y_i), i = 1,...N\)
were generated for \(x_i \in [x_{min}, x_{max}]\) where \(x\) and \(y\)
have a linear relationship. Data were simulated based on linear model
with additive errors: \begin{align}
y_i & = \beta_0 + \beta_1 x_i + e_i \\
\text{with } e_i & \sim N(0, \sigma^2). \nonumber
\end{align} The parameters \(\beta_0\) and \(\beta_1\) are selected to
replicate Mosteller et al.~(1981) with \(e_i\) generated by rejection
sampling in order to guarantee the points shown align with that of the
fitted line. An ordinary least squares regression is then fit to the
simulated points in order to obtain the best fit line and fitted values
in 0.25 increments across the domain,
\((x_k, \hat y_{k,OLS}), k = 1, ..., 4 x_{max} +1\). The data simulation
function then outputs a list of point data and line data both indicating
the parameter identification, x-value, and corresponding simulated or
fitted y value. The data simulation procedure is described in
\cref{alg:eyefitting-algorithm}.

\begin{algorithm}
  \caption{Eye Fitting Straight Lines in the Modern Era Data Simulation}\label{alg:eyefitting-algorithm}
  \begin{algorithmic}[1]
    \Statex \textbullet~\textbf{Input Parameters:} $y_{\bar{x}}$ for calculating the y-intercept, $\beta_0$; slope $\beta_1$; standard deviation from line $\sigma$; sample size of points $N = 30$; domain $x_{min}$ and $x_{max}$; fitted value increment $x_{by} = 0.25$.
    \Statex \textbullet~\textbf{Output Parameters:} List of point data and line data each indicating the parameter identification, x value, and corresponding simulated or fitted y value.
    \State Randomly select and jitter $N = 30$ x-values along the domain, $x_{i=1:N}\in [x_{min}, x_{max}]$.
    \State Determine the y-intercept, $\beta_0$, at x = 0 from the provided slope ($\beta_1$) and y-value at the mean of x ($y_{\bar{x}}$) using point-slope equation of a line.
    \State Generate "good" errors, $e_{i = 1:N}$ based on $N(0,\sigma)$ by setting a constraint requiring the mean of the first $\frac{1}{3}\text{N}$ errors $< |2\sigma|.$
    \State Simulate point data based on $y_i = \beta_0 + \beta_1 x_i + e_i$
    \State Obtain ordinary least squares regression coefficients, $\hat\beta_0$ and $\hat\beta_1$, for the simulated point data using the lm function in the stats package in base R.
    \State Obtain fitted values every 0.25 increment across the domain from the ordinary least squares regression $\hat y_{k,OLS} = \hat\beta_{0,OLS} + \hat\beta_{1,OLS} x_k$.
    \State Output data list of point data and line data each indicating the parameter identification, x value, and corresponding simulated or fitted y value.
  \end{algorithmic}
\end{algorithm}

@@

Simulated model equation parameters were selected to reflect the four
data sets (F, N, S, and V) used in \citet{mosteller_eye_1981}
(\cref{tab:eyefitting-parameters}). Parameter choices F, N, and S
simulated data across a domain of 0 to 20. Parameter choice F produces a
trend with a positive slope and a large variance while N has a negative
slope and a large variance. In comparison, S shows a trend with a
positive slope with a small variance and V yields a steep positive slope
with a small variance over the domain of 4 to 16.
\cref{fig:eyefitting-simplot} illustrates an example of simulated data
for all four parameter choices intended to reflect the trends seen in
\cref{fig:mosteller-eyefitting-plot}. Aesthetic design choices were made
consistent across each of the interactive you draw it plots; the aspect
ratio, defining the \(x\) to \(y\) axis ratio was set to one and the
y-axis range extended 10\% beyond (above and below) the range of the
simulated data points to allow for users to draw outside the simulated
data set range.

@@

{\centering \includegraphics[width=0.75\linewidth,]{Eye-Fitting-Stright-Lines-in-the-Modern-Era_files/figure-latex/eyefitting-simplot-1}

@@

Participants completed the experiment using a RShiny application found
\href{https://shiny.srvanderplas.com/you-draw-it/}{here}. During May
2021, participants were recruited through Twitter, Reddit, and direct
email. A total of 39 individuals completed 256 unique you draw it task
plots; all completed you draw it task plots were included in the
analysis.

@@

@@

{\centering \includegraphics[width=1\linewidth,]{Eye-Fitting-Stright-Lines-in-the-Modern-Era_files/figure-latex/ols-vs-pca-example-1}

\caption{OLS vs PCA Regression Lines}\label{fig:ols-vs-pca-example}
\end{figure}

For each participant, the final data set used for analysis contains
\(x_{ijk}, y_{ijk,drawn}, \hat y_{ijk,OLS}\), and \(\hat y_{ijk,PCA}\)
for parameter choice \(i = 1,2,3,4\), j = \(1,...N_{participant}\), and
\(x_{ijk}\) value \(k = 1, ...,4 x_{max} + 1\). Using both a linear
mixed model and a generalized additive mixed model, comparisons of
vertical residuals in relation to the OLS fitted values
(\(e_{ijk,OLS} = y_{ijk,drawn} - \hat y_{ijk,OLS}\)) and PCA fitted
values (\(e_{ijk,PCA} = y_{ijk,drawn} - \hat y_{ijk,PCA}\)) were made
across the domain. \cref{fig:eyefitting-example-plot} displays an
example of all three fitted trend lines for parameter choice F.

{\centering \includegraphics[width=0.65\linewidth,]{Eye-Fitting-Stright-Lines-in-the-Modern-Era_files/figure-latex/eyefitting-example-plot-1}

\caption{Eye Fitting Straight Lines in the Modern Era Example}\label{fig:eyefitting-example-plot}
\end{figure}

Using the lmer function in the lme4 package \citep{lme4}, a linear mixed
model (LMM) is fit separately to the OLS and PCA residuals, constraining
the fit to a linear trend. Parameter choice, \(x\), and the interaction
between \(x\) and parameter choice were treated as fixed effects with a
random participant effect accounting for variation due to participant.
The LMM equation for each fit (OLS and PCA) residuals is given by:
\begin{equation}
y_{ijk,drawn} - \hat y_{ijk,fit} = e_{ijk,fit} = \left[\gamma_0 + \alpha_i\right] + \left[\gamma_{1} x_{ijk} + \gamma_{2i} x_{ijk}\right] + p_{j} + \epsilon_{ijk}
\end{equation} \noindent where

\begin{itemize}
\tightlist
\item
  \(y_{ijk,drawn}\) is the drawn y-value for the \(i^{th}\) parameter
  choice, \(j^{th}\) participant, and \(k^{th}\) increment of x-value
\item
  \(\hat y_{ijk,fit}\) is the fitted y-value for the \(i^{th}\)
  parameter choice, \(j^{th}\) participant, and \(k^{th}\) increment of
  x-value corresponding to either the OLS or PCA fit
\item
  \(e_{ijk,fit}\) is the residual between the drawn and fitted y-values
  for the \(i^{th}\) parameter choice, \(j^{th}\) participant, and
  \(k^{th}\) increment of x-value corresponding to either the OLS or PCA
  fit
\item
  \(\gamma_0\) is the overall intercept
\item
  \(\alpha_i\) is the effect of the \(i^{th}\) parameter choice (F, S,
  V, N) on the intercept
\item
  \(\gamma_1\) is the overall slope for \(x\)
\item
  \(\gamma_{2i}\) is the effect of the parameter choice on the slope
\item
  \(x_{ijk}\) is the x-value for the \(i^{th}\) parameter choice,
  \(j^{th}\) participant, and \(k^{th}\) increment
\item
  \(p_{j} \sim N(0, \sigma^2_{participant})\) is the random error due to
  the \(j^{th}\) participant's characteristics
\item
  \(\epsilon_{ijk} \sim N(0, \sigma^2)\) is the residual error.
\end{itemize}

Eliminating the linear trend constraint, the bam function in the mgcv
package \citep{mgcv1, mgcv2, mgcv3, mgcv4, mgcv5} is used to fit a
generalized additive mixed model (GAMM) separately to the OLS and PCA
residuals to allow for estimation of smoothing splines. Parameter choice
was treated as a fixed effect with no estimated intercept and a separate
smoothing spline for \(x\) was estimated for each parameter choice. A
random participant effect accounting for variation due to participant
and a random spline for each participant accounted for variation in
spline for each participant. The GAMM equation for each fit (OLS and
PCA) residuals is given by: \begin{equation}
y_{ijk, drawn} - \hat y_{ijk, fit} = e_{ijk,fit} = \alpha_i + s_{i}(x_{ijk}) + p_{j} + s_{j}(x_{ijk})
\end{equation} \noindent where

\begin{itemize}
\tightlist
\item
  \(y_{ijk,drawn}\) is the drawn y-value for the \(i^{th}\) parameter
  choice, \(j^{th}\) participant, and \(k^{th}\) increment of x-value
\item
  \(\hat y_{ijk,fit}\) is the fitted y-value for the \(i^{th}\)
  parameter choice, \(j^{th}\) participant, and \(k^{th}\) increment of
  x-value corresponding to either the OLS or PCA fit
\item
  \(e_{ijk,fit}\) is the residual between the drawn and fitted y-values
  for the \(i^{th}\) parameter choice, \(j^{th}\) participant, and
  \(k^{th}\) increment of x-value corresponding to either the OLS or PCA
  fit
\item
  \(\alpha_i\) is the intercept for the parameter choice \(i\)
\item
  \(s_{i}\) is the smoothing spline for the \(i^{th}\) parameter choice
\item
  \(x_{ijk}\) is the x-value for the \(i^{th}\) parameter choice,
  \(j^{th}\) participant, and \(k^{th}\) increment
\item
  \(p_{j} \sim N(0, \sigma^2_{participant})\) is the error due to
  participant variation
\item
  \(s_{j}\) is the random smoothing spline for each participant.
\end{itemize}

\cref{fig:eyefitting-lmer-residualplots} and
\cref{fig:eyefitting-gamm-residualplots} show the estimated trends of
residuals (vertical deviation of participant drawn points from both the
OLS and PCA fitted points) as modeled by a LMM and GAMM respectively.
Examining the plots, the estimated trends of PCA residuals (orange)
appear to align closer to the \(y=0\) horizontal (dashed) line than the
OLS residuals (blue). In particular, this trend is more prominent in
parameter choices with large variances (F and N). These results are
consistent to those found in \citet{mosteller_eye_1981} indicating
participants fit a trend line closer to the estimated regression line
with the slope of the first principal component than the estimated OLS
regression line.

{\centering \includegraphics[width=0.9\linewidth,]{Eye-Fitting-Stright-Lines-in-the-Modern-Era_files/figure-latex/eyefitting-lmer-residualplots-1}

\caption{Eye Fitting Straight Lines in the Modern Era LMM results}\label{fig:eyefitting-lmer-residualplots}
\end{figure}

{\centering \includegraphics[width=0.9\linewidth,]{Eye-Fitting-Stright-Lines-in-the-Modern-Era_files/figure-latex/eyefitting-gamm-residualplots-1}

\caption{Eye Fitting Straight Lines in the Modern Era GAMM results}\label{fig:eyefitting-gamm-residualplots}
\end{figure}

@@

The intent of this paper was to establish you draw it as a tool. Eye
Fitting Straight Lines in the Modern Era replicated the results found in
Mosteller et al.~(1981). When shown points following a linear trend,
participants tended to fit the slope of the first principal component
over the slope of the least squares regression line. This trend was most
prominent when shown data simulated with larger variances. The
reproducibility of these results serve as evidence of the reliability of
the you draw it method.

@@

Use the method. \emph{Should probably elaborate here..}

>>>>

\usepackage[dvipsnames]{xcolor} % colors
\newcommand{\ear}[1]{{\textcolor{blue}{#1}}}
\newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
\newcommand{\rh}[1]{{\textcolor{Green}{#1}}}
\usepackage[capitalise]{cleveref}
\newcommand\pcref[1]{(\cref{#1})}
\usepackage{algorithm,algpseudocode,booktabs}

@@

\begin{itemize}
\tightlist
\item
  What are graphs? Why do we care?
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Cleveland and McGill
\end{itemize}

@@

Graphical tests are useful for studying the perception of statistical
graphs. Studies might ask participants to identify differences in
graphs, read information off of a chart accurately, use data to make
correct real-world decisions, or predict the next few observations. All
of these types of tests require different levels of use and manipulation
of the information being presented in the chart. Early researchers
studied graphs from a psychological perspective. These studies generally
tested participants ability to detect a stimulus (or a difference
between two stimuli) ({\textcolor{blue}{CITATIONS}}). Here we focus on
the how graphical testing has developed in statistics.

A major development in statistical graphics research is Wilkinson's
Grammar of Graphics \citep{wilkinson2013grammar}. The grammar of
graphics serves as the fundamental framework for data visualization with
the notion that graphics are built from the ground up by specifying
exactly how to create a particular graph from a given data set. Visual
representations are constructed through the use of ``tidy data'' which
is characterized as a data set in which each variable is in its own
column, each observation is in its own row, and each value is in its own
cell \citep{wickham2016r}. Graphics are viewed as a mapping from
variables in a data set (or statistics computed from the data) to visual
attributes such as the axes, colors, shapes, or facets on the canvas in
which the chart is displayed. Software, such as Hadley Wickham's ggplot2
\citep{wickham2011ggplot2}, aims to implement the framework of creating
charts and graphics as the grammar of graphics recommends.

One useful tool for testing statistical graphics is the concept of a
lineup. \citet{buja2009statistical} introduced the lineup protocol in
which data plots are depicted and interpreted as statistics. Supported
by the grammar of graphics, a data plot can be characterized as a
statistic, defined as, ``a functional mapping of a variable or set of
variables'' \citep{vanderplas2020testing}. This allows the data plot to
be tested similar to other statistics, by comparing the actual data plot
to a set of plots with the absence of any data structure we can test the
likelihood of any perceived structure being significant. The
construction of data plots as statistics allow for easy experimentation,
granting researchers the ability to compare the effectiveness of and
understand the perception of different types of charts
({\textcolor{blue}{CITATIONS}}). The lineup protocol is one such example
of the development of tools designed for statistical graphical testing.
The advancement of graphing software provides the tools necessary to
develop new methods of testing graphics.

\hypertarget{fitting-trends-by-eye}{%
\subsection{Fitting Trends by Eye}\label{fitting-trends-by-eye}}

Initial studies in the 20th century explored the use of fitting lines by
eye through a set of points
\citep{finney1951subjective, mosteller1981eye}. Common methods of
fitting trends by eye involved maneuvering a string, black thread, or
ruler until the fit is suitable, then drawing the line through the set
of points. Recently, \citet{ciccione2021can} conducted a comprehensive
set of studies investigating human ability to detect trends in graphical
representations from a psychophysical approach.

In \citet{finney1951subjective}, it was of interest to determine the
effect of stopping iterative maximum likelihood calculations after one
iteration. Many techniques in statistical analysis are performed with
the aid of iterative calculations such as Newton's method or Fisher's
scoring. The author was interested in whether one iteration of
calculations was sufficient in the estimation of parameters connected
with dose-response relationships. One measure of interest is the
relative potency between a test preparation of doses and standard
preparation of does; relative potency is calculated as the ratio of two
equally effective doses between the two preparation methods. In this
study, twenty-one scientists were recruited via postal mail and asked to
``rule two lines'' in order to judge by eye the positions for a pair of
parallel probit regression lines in a biological assay. The author then
computed one iterative calculation of the relative potency based on
starting values as indicated by the pair of lines provided by each
participant and compared these relative potency estimates to that which
was estimated by the full probit technique (reaching convergence through
multiple iterations). Results indicated that one cycle of iterations for
calculating the relative potency was sufficient based on the starting
values provided by eye from the participants.

\citet{mosteller1981eye}, sought to understand the properties of least
squares and other computed lines by establishing one systematic method
of fitting lines by eye. Participants were asked to fit lines by eye to
four scatter-plots using an 8.5 x 11 inch transparency with a straight
line etched completely across the middle. A latin square design with
packets of the set of points stapled together in four different
sequences was used to determine if there is an effect of order of
presentation. It was found that order of presentation had no effect and
that participants tended to fit the slope of the principal axis (error
minimized orthogonally, both horizontal and vertical, to the regression
line) over the slope of the least squares regression line (error
minimized vertically to the regression line). These results support
previous research on ``ensemble perception'' indicating the visual
system can compute averages of various features in parallel across the
items in a set ({\textcolor{blue}{CITATIONS}}).

In \citet{ciccione2021can}, participants were asked to judge trends,
estimate slopes, and conduct extrapolation. To estimate slopes,
participants were asked to report the slope of the best-fitting
regression line using a trackpad to adjust the tilt of a line on screen.
Results indicated the slopes participants reported were always in excess
of the ideal slopes, both in the positive and in the negative direction,
and those biases increase with noise and with number of points. This
supports the results found in \citet{mosteller1981eye} and suggest that
participants might use Deming regression when fitting a line to a noisy
scatterplot.

In 2015, the New York Times introduced an interactive feature, called
You Draw It
\citep{aisch_cox_quealy_2015, buchanan_park_pearce_2017, katz_2017}.
Readers are asked to input their own assumptions about various metrics
and compare how these assumptions relate to reality. The New York Times
team utilizes Data Driven Documents (D3) that allows readers to predict
these metrics through the use of drawing a line on their computer screen
with their computer mouse. After the reader has completed drawing the
line, the actual observed values are revealed and the reader may check
their estimated knowledge against the actual reported data.

\hypertarget{research-objectives}{%
\subsection{Research objectives}\label{research-objectives}}

In this paper, we establish `You Draw It', adapted from the New York
Times feature, as tool for graphical testing. The `You Draw It' method
is validated by replicating the study conducted by
\citet{mosteller1981eye}. The results of our study provide support for
``ensemble perception''.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

Participants were recruited through through Twitter, Reddit, and direct
email in May 2021. A total of 39 individuals completed 256 unique `You
Draw It' task plots; all completed you draw it task plots were included
in the analysis. All participants had normal or corrected to normal
vision and signed an informed consent form. The experimental tasks took
approximately 15 minutes to complete. Participants completed the
experiment on their own computers in an environment of their choosing.
The experiment was conducted and distributed through an RShiny
application found
\href{https://shiny.srvanderplas.com/you-draw-it/}{here}.

\hypertarget{you-draw-it-task}{%
\subsection{`You Draw It' Task}\label{you-draw-it-task}}

Data Driven Documents (D3), a JavaScript-based graphing framework that
facilitates user interaction, is used to create the `You Draw It' task
plots. Integrating this into RShiny using the r2d3 package, participants
are asked to draw a line using their computer mouse through trend shown
on their screen. Details of the development of the `You Draw It' task
plots will be addressed in future work. In the study, participants are
shown an interactive scatter-plot \cref{fig:ydi-stimuli} along with the
prompt, ``Use your mouse to fill in the trend in the yellow box
region.'' The yellow-box region moves along as the participant draws
their trend-line until the yellow region disappears, indicating the
participant has filled in the entire domain.

{\centering \includegraphics[width=1\linewidth,]{images/ydi-stimuli}

\caption{'You Draw It' Stimuli}\label{fig:ydi-stimuli}
\end{figure}

@@

All data processing was conducted in R statistical software. A total of
\(N = 30\) points \((x_i, y_i), i = 1,...N\) were generated for
\(x_i \in [x_{min}, x_{max}]\) where \(x\) and \(y\) have a linear
relationship. Data were simulated based on linear model with additive
errors: \begin{align}
y_i & = \beta_0 + \beta_1 x_i + e_i \\
\text{with } e_i & \sim N(0, \sigma^2). \nonumber
\end{align} The parameters \(\beta_0\) and \(\beta_1\) are selected to
replicate \citet{mosteller1981eye} with \(e_i\) generated by rejection
sampling in order to guarantee the points shown align with that of the
fitted line.

Simulated model equation parameters were selected to reflect the four
data sets (F, N, S, and V) used in \citet{mosteller1981eye}
(\cref{tab:eyefitting-parameters}). Parameter choices F, N, and S
simulated data across a domain of 0 to 20. Parameter choice F produces a
trend with a positive slope and a large variance while N has a negative
slope and a large variance. In comparison, S shows a trend with a
positive slope with a small variance and V yields a steep positive slope
with a small variance over the domain of 4 to 16.
\cref{fig:eyefitting-simplot} illustrates an example of simulated data
for all four parameter choices intended to reflect the trends in
\citet{mosteller1981eye}. Aesthetic design choices were made consistent
across each of the interactive you draw it plots. The y-axis range
extended 10\% beyond (above and below) the range of the simulated data
points to allow for users to draw outside the simulated data set range.

@@

@@

{\centering \includegraphics[width=1\linewidth,]{Eye-Fitting-Stright-Lines-in-the-Modern-Era_files/figure-latex/eyefitting-simplot-1}

@@

This experiment was conducted as part of a larger study; for simplicity,
we focus on the study design and methods related to the current study.
Each scatter-plot was the graphical representation of a data set that
was generated randomly, independently for each participant at the start
of the experiment. Participants in the study are shown two `You Draw It'
practice plots in order to train participants in the skills associated
with executing the task followed by four task plots associated with the
current study. The order of the task plots was randomly assigned for
each individual in a completely randomized design.

@@

\hypertarget{fitted-values}{%
\subsection{Fitted Values}\label{fitted-values}}

@@

\begin{itemize}
\tightlist
\item
  Because of the randomness in the stimulus generation process, on any
  given trial the actual slope of the linear regression line could
  differ from its prescribed value.
\end{itemize}

\hypertarget{feedback-data}{%
\subsection{Feedback Data}\label{feedback-data}}

\hypertarget{linear-trend-constraint-performance}{%
\subsection{Linear Trend Constraint
Performance}\label{linear-trend-constraint-performance}}

\hypertarget{smoothing-spline-trend-performance}{%
\subsection{Smoothing Spline Trend
Performance}\label{smoothing-spline-trend-performance}}

@@

@@

\begin{itemize}
\tightlist
\item
  Use the method.
\item
  Write R package.
\end{itemize}
